\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-09 paper.tex
small changes

Automatically inducing programs from examples is a long-standing goal
of artificial intelligence. Recent work has successfully used
symbolic search techniques (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}), neural networks trained from
a corpus of examples (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}) to
synthesize programs for task domains such as string transformations,
list processing, and robot navigation and planning. However, all these
approaches -- symbolic, neural and neural-symbolic -- rely upon a
hand-engineered \emph{Domain-Specific Language} (DSL). The DSL is an
inventory of restricted programming primitives, encoding domain-specific
knowledge about the space of programs. In
practice
we often have only a few input/output examples for each
program to be induced, and thus success often hinges on having a good
DSL that provides a crucial inductive bias for what would otherwise be an
unconstrained search through the space of all computable functions.
Here we ask, to what extent can we dispense with such highly
hand-engineered domain-specific languages?

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference in the generative model of Fig.~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability,
alternating maximization w.r.t. the frontiers (Wake) and the DSL (Sleep-G):
%So interleaved with the DSL induction steps are program induction steps. Sec.~\ref{explorationSection}
%explains how \system synthesizes new programs and given a DSL.
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x,p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t. the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

@@

Now our goal is to search for programs that solve the tasks.  In this
work we use the simple search strategy of enumerating programs from
the DSL  in decreasing order of their probability,
and then checking if an enumerated program $p$ assigns positive
probability to a task ($\probability[x|p] > 0$); if so, we include $p$ in
the frontier $\mathcal{F}_x$.

>>>>

Automatically inducing programs from examples is a long-standing goal
of artificial intelligence. Recent work has successfully used
symbolic search techniques (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}), neural networks trained from
a corpus of examples (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}) to
synthesize programs for task domains such as string transformations,
list processing, and robot navigation and planning. However, all these
approaches -- symbolic, neural and neural-symbolic -- rely upon a
hand-engineered \emph{Domain-Specific Language} (DSL). The DSL is an
inventory of restricted programming primitives, encoding domain-specific
knowledge about the space of programs. In
practice
we often have only a few input/output examples for each
program to be induced, and thus success often hinges on having a good
DSL that provides a crucial inductive bias for what would otherwise be an
unconstrained search through the space of all computable functions.
Here we ask, to what extent can we dispense with such highly
hand-engineered domain-specific languages?

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference in the generative model of Fig.~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability,
alternating maximization w.r.t. the frontiers (Wake) and the DSL (Sleep-G):
%So interleaved with the DSL induction steps are program induction steps. Sec.~\ref{explorationSection}
%explains how \system synthesizes new programs and given a DSL.
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x,p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t. the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

@@

Now our goal is to search for programs that solve the tasks.  In this
work we use the simple search strategy of enumerating programs from
the DSL  in decreasing order of their probability,
and then checking if an enumerated program $p$ assigns positive
probability to a task ($\probability[x|p] > 0$); if so, we incorporate $p$ 
into the frontier $\mathcal{F}_x$.
